Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-03-10 paper.qmd
Removed MNIST third projection

distance_df_small_edges_mnist <- distance_mnist |>
  filter(distance < benchmark_mnist) #|>
  #filter(!(row_number() %in% c(155, 554, 140)))

@@

# trimesh_removed_mnist <- vis_rmlg_mesh(
#   distance_edges = distance_mnist, 
#   benchmark_value = benchmark_mnist, 
#   tr_coord_df = tr_from_to_df_mnist, 
#   distance_col = "distance") +
#   geom_point(data = pacmap_minst_scaled,
#              aes(
#                x = PaCMAP1,
#                y = PaCMAP2
#              ),
#              alpha=0.5, 
#              size = 0.1) +
#   interior_annotation("a", sc_ltr_pos_mnist)

@@

# error_df <- error_df |>
#   mutate(type = case_when(
#     row_wise_abs_error <= 2 ~ "error 0-2",
#     row_wise_abs_error <= 4 ~ "error 2-4",
#     row_wise_abs_error <= 6 ~ "error 4-6",
#     row_wise_abs_error <= 8 ~ "error 6-8",
#     row_wise_abs_error <= 10 ~ "error 8-10",
#     .default = "error greter than 10"
#   )) |>
#   mutate(type = factor(type, levels = c(
#     "error 0-2", "error 2-4", "error 4-6", "error 6-8", "error 8-10", 
#     "error greter than 10")))

@@

highd_error_2d_mnist <- error_df |>
  ggplot(aes(x = tSNE1,
             y = tSNE2,
             colour = sqrt_row_wise_tot_error)) +
  geom_point(alpha=0.2) +
  scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
  geom_rect(aes(xmin = 0.82, xmax = 0.92, ymin = 0.38, ymax = 0.58),
               fill = "transparent", color = "#000000", linewidth = 0.7) +
  geom_rect(aes(xmin = 0.9, xmax = 1, ymin = 1.25, ymax = 1.45),
               fill = "transparent", color = "#e41a1c", linewidth = 0.7)  +
  geom_rect(aes(xmin = 0.35, xmax = 0.45, ymin = 0.55, ymax = 0.7),
               fill = "transparent", color = "#4daf4a", linewidth = 0.7)  +
  geom_rect(aes(xmin = 0.7, xmax = 0.8, ymin = 0.25, ymax = 0.4),
               fill = "transparent", color = "#984ea3", linewidth = 0.7) +
  theme(legend.position = "none")
  # scale_colour_brewer(
  #   type = "seq",
  #   palette = 1
  # ) #+ 
  # theme(
  #     legend.position = "bottom",
  #     legend.text = element_text(size = 10),          # Increase legend text size
  #     legend.key.size = unit(1, "lines")              # Increase size of legend keys
  # ) +
  # guides(
  #     color = guide_legend(
  #         override.aes = list(size = 2.5, alpha = 1)      # Increase point size and opacity in legend
  #     )
  # ) 
```

@@

scaled_mnist_data <- scaled_mnist |>
  filter(type == "data") |>
  # bind_cols(tsne_minst_scaled_temp) |>
  # select(-type, -tSNE1, -tSNE2, -ID) |>
  select(-type)

# scaled_mnist_data <- bind_cols(scaled_mnist_data,
#                                tsne_minst_scaled |> select(-ID))

@@

# model_2d <- df_bin_centroids_mnist |>
#   select(c_x, c_y) |>
#   rename(c("tSNE1" = "c_x",
#            "tSNE2" = "c_y"))
# 
# scaled_mnist_data_model <- bind_cols(scaled_mnist_data_model, model_2d)

# Combine with the true model for visualization
# df <- dplyr::bind_rows(scaled_mnist_data_model |> mutate(type = "model"),
#                        scaled_mnist_data)

# scaled_mnist_data <- scaled_mnist_data |>
#   select(-type)

## First projection
# projection <- cbind(
#     c(0.71503,0.00502,-0.04370,-0.01811,-0.01103,0.09877,0.06199,0.05396,0.07771,-0.01713),
#     c(0.00587,0.71879,0.03485,0.08284,-0.07534,-0.05420,0.04315,-0.00547,-0.02995,0.01810))

@@

mnist_proj_tsne_model2_link <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2,
      colour = cluster)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    linewidth = 0.5) +
  geom_point(
    size = 1.2,
    alpha = 0.3) +
  scale_color_manual(values = c("#d9d9d9",'#ff7f00')) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 2) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("c", cex = 2) +
  theme(
    legend.position = "none"
  )

## Third projection
projection <- cbind(
    c(-0.12783,-0.50063,-0.02631,-0.00484,0.17733,0.37157,0.11310,0.09424,0.19844,0.19792),
    c(-0.42542,0.29570,0.31106,-0.30307,0.00614,0.16714,0.21779,-0.00025,0.07020,-0.00689))

projection_scaled <- projection * 1.5

mnist_proj_tsne_model3 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2,
      colour = cluster)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") +
  geom_point(
    #size = 0.2,
    alpha = 0.3) +
  scale_color_manual(values = c("#999999",'#ff7f00')) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 2) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 0.6)) +
  ylim(c(-1, 0.6)) +
  interior_annotation("d", cex = 1.5) +
  theme(
    legend.position = "none"
  )

@@

free(trimesh_removed_mnist) + mnist_proj_tsne_model1 +
  #mnist_proj_tsne_model2 + mnist_proj_tsne_model3 +
  mnist_proj_tsne_model2 +
  plot_layout(ncol = 3)

>>>>

distance_df_small_edges_mnist <- distance_mnist |>
  filter(distance < benchmark_mnist)

@@

@@

@@

highd_error_2d_mnist <- error_df |>
  ggplot(aes(x = tSNE1,
             y = tSNE2,
             colour = sqrt_row_wise_tot_error)) +
  geom_point(alpha=0.2) +
  scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
  geom_rect(aes(xmin = 0.82, xmax = 0.92, ymin = 0.38, ymax = 0.58),
               fill = "transparent", color = "#000000", linewidth = 0.7) +
  geom_rect(aes(xmin = 0.9, xmax = 1, ymin = 1.25, ymax = 1.45),
               fill = "transparent", color = "#e41a1c", linewidth = 0.7)  +
  geom_rect(aes(xmin = 0.35, xmax = 0.45, ymin = 0.55, ymax = 0.7),
               fill = "transparent", color = "#4daf4a", linewidth = 0.7)  +
  geom_rect(aes(xmin = 0.7, xmax = 0.8, ymin = 0.25, ymax = 0.4),
               fill = "transparent", color = "#984ea3", linewidth = 0.7) +
  theme(legend.position = "none")
```

@@

scaled_mnist_data <- scaled_mnist |>
  filter(type == "data") |>
  select(-type)

@@

@@

mnist_proj_tsne_model2_link <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2,
      colour = cluster)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    linewidth = 0.5) +
  geom_point(
    size = 1.2,
    alpha = 0.3) +
  scale_color_manual(values = c("#d9d9d9",'#ff7f00')) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 2) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("c", cex = 2) +
  theme(
    legend.position = "none"
  )
```

@@

free(trimesh_removed_mnist) + mnist_proj_tsne_model1 +
  mnist_proj_tsne_model2 +
  plot_layout(ncol = 3)
